Tidal Heating via Internal Gravity Waves in White Dwarf Binaries

### 2019-02-04 writeups/paper.tex
stash paper

    \documentclass[twocolumn,
        nofootinbib,
        usenames, % allows access to some tikz colors
        dvipsnames % more colors: https://en.wikibooks.org/wiki/LaTeX/Colors
    ]{revtex4-1}% chktex 8
    \usepackage{
        amsmath,
        amssymb,
        fancyhdr, % page styling
        lastpage, % footer fanciness
        hyperref, % various links
        setspace, % line spacing
        amsthm, % newtheorem and proof environment
        mathtools, % \Aboxed for boxing inside aligns, among others
        float, % Allow [H] figure env alignment
        enumerate, % Allow custom enumerate numbering
        graphicx, % allow includegraphics with more filetypes
        wasysym, % \smiley!
        upgreek, % \upmu for \mum macro
        listings, % writing TrueType fonts and including code prettily
        tikz, % drawing things
        booktabs, % \bottomrule instead of hline apparently
        cancel % can cancel things out!
    }
    \usepackage[
        labelfont=bf, % caption names are labeled in bold
        font=scriptsize % smaller font for captions
    ]{caption}
    \usepackage[font=scriptsize]{subcaption} % subfigures

@@

In \autoref{s:num_setup}, we will cover the numerical setup. In \autoref{s:lin}, we
will discuss the agreement of our simulation in the low-amplitude limit with
analytical linear theory. In \autoref{s:nonlin} we will present a few reference
simulations illustrating critical layer absorption depositing horizontal angular
momentum into the fluid.

\section{Problem Description}\label{s:num_setup}

\subsection{Fluid Equations}

We consider an incompressible, isothermal fluid, representative of degenerate
matter in WD bulks, in 2D. We assume a barotropic equation of state as a first
approximation. As we are interested in the behavior of the wave far from the
center of the WD, we approximate the gravitational field as uniform. We model
the background density stratification as $\rho_0(x, z) = \rho_0(z) = \rho_{ref}
e^{-z/H}$ for some reference density $\rho_{ref} = \at{\rho_0(z)}_{z = 0}$.

@@

\subsection{Internal Gravity Waves: Linear Theory}

In the small perturbation limit, where flow velocities are small compared to the
characteristic space and time scales $\pd{}{t} \gg \vec{u} \cdot \vec{\nabla}$,
we linearize \autoref{se:fc_orig} by ignoring the advective components of the
material derivative, defining $\rho_1 \equiv \rho - \rho_0 \ll \rho_0$ and $P_1
\equiv P - P_0$ deviations from equilibrium and obtain:
\begin{subequations}\label{se:lin_homo}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} &= 0,\\
        \pd{\rho_1}{t} - \frac{u_{z} \rho_0}{H} &= 0,\\
        \pd{\vec{u}}{t} + \frac{\vec{\nabla}P_1}{\rho_0}
            + \frac{\rho_1 g\hat{z}}{\rho_0}
            &= 0.
    \end{align}
\end{subequations}

It is well known that the solutions to the $F = 0$ homogeneous
\autoref{se:lin_homo} are of form (cite)
\begin{subequations}\label{se:lin_sol}
    \begin{align}
        u_z\p*{x, z, t} &= Ae^{z/2H}e^{i(k_{0x}x + k_{0z}z - \omega_0 t)},\\
        \omega_0^2 &= \frac{N^2k_{0x}^2}{k_{0x}^2 + k_{0z}^2 + \frac{1}{4H^2}},
    \end{align}
\end{subequations}
up to undetermined amplitude $A$, where
\begin{equation}
    N^2 \equiv g^2\p*{\rd{\rho}{P} - \at{\pd{\rho}{P}}_{ad}}
        = \frac{g}{H},
\end{equation}
is the Brunt-V\"ais\"al\"a, or buoyancy, frequency.

In the weak stratification limit $k_zH \gg 1$, the solution exhibits the
following characteristics:
\begin{itemize}
    \item The amplitude of the perturbations grows like $e^{z/2H}$. Thus, the
        linear approximation is violated for sufficiently large $z$.

    \item The phase and group velocities can be computed respectively:
        \begin{align}
            \vec{c}_{ph} &= \frac{Nk_x}{k_x^2 + k_z^2 + 1/4H^2} \hat{k},\\
            \vec{c}_{g} &= \frac{N\p*{k_xk_z\hat{x}
                - \p*{k_z^2 + \frac{1}{4H^2}}\hat{z}}}
                {\p*{k_x^2 + k_z^2 + \frac{1}{4H^2}}^{3/2}} \nonumber\\
                &\approx \frac{Nk_z}{k_x^2 + k_z^2}\p*{k_x\hat{x} - k_z\hat{z}}.
        \end{align}
        We recover the usual result $\vec{c}_{ph} \cdot \vec{c}_g = 0$ (CITE).
        Note that an IGW transporting energy and momentum upwards has $k_z < 0$.

%     \item The linear solution is an exact solution to the full nonlinear
%         equations \autoref{se:fc_orig}, thanks to the advective terms vanishing
%         identically when $\vec{\nabla} \cdot \vec{u} = 0$ (CITE). However, this
%         solution is globally unstable to resonant triad interactions (CITE).

    \item The time-averaged total $x$-momentum flux in the $\hat{z}$ direction
        can be computed. Since the fluid variables have separable as $f(x, z, t)
        = f(z)e^{i(k_{0x}x - \omega_0 t)}$, $x$ averaging and time averaging are
        equivalent, so we may write ($\ev*{\dots}_x$ denotes $x$-averaging)
        \begin{align}
            S_{px, z} &\equiv \ev*{\rho u_x u_z}_x =
                \frac{1}{L_x}\int_0^{L_x}\limits \rho u_x u_z\;\mathrm{d}x
                    \nonumber\\
                &\approx -\frac{A^2}{2}\rho_{ref}\frac{k_z}{k_x}.
                    \label{eq:fpx_lin}
        \end{align}

    \item Finally, it may be noted that horizontally averaged horizontal
        momentum $p_x = \ev*{\rho u_x}_x$ obeys $\pd{p_x}{t} + \pd{S_{px}}{z} =
        0$. Since flux is transported on the group velocity timescale (CITE),
        this simplifies to the known expression for the wave-induced horizontal
        mean flow (cite)
        \begin{equation}
            \bar{U}_0 \equiv \ev*{u_x}_x = \frac{\ev*{u_xu_z}_x}{c_{gz}}
        \end{equation}
        We've denoted $c_{gz} = \at{\pd{\omega}{k_{z}}}_{k_{z0}}$ the vertical group
        velocity. In terms of our forcing amplitude,
        \begin{equation}
            \bar{U}_0(F) = -\frac{A(F)^2}{2}e^{\frac{z - z_0}{H}} \frac{k_z}{k_x}c_{gz},
                \label{eq:u0_lin}
        \end{equation}
        where $c_{gz} = -\frac{Nk_{0x}k_{0z}}{\p*{k_{0x}^2 + k_{0z}^2 +
        \frac{1}{4H^2}}^{3/2}}$.
\end{itemize}
% Linear solution: e^z/2H, dispersion relation, phase/group velocity. Observe is
% exact solution of nonlinear.
% Linear solution is globally unstable to RTI though, thus not stable
% equilibrium of IVP
%
% Nonlinear: Critical layer absorption (viscous = Hazel, nonlin = Winters),
% wave-induced mean flow, horizontal momentum flux

@@

As not to interfere with the incompressibility constraint, we force the system
on the density equation, replacing \autoref{eq:fc_density} with
\begin{equation}
    \md{\rho}{t} = Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
            \cos \p*{k_{0x}x - \omega_0 t}.
\end{equation}
Using a narrow Gaussian profile excites a broad $k_z$ wavenumber spectrum, and
only the $k_{0z}$ satisfying the IGW's dispersion relation for the given
$k_{0x}, \omega_0(k_{0x}, k_{0z})$ will propagate. The fluid equations then take
form
\begin{subequations}\label{se:vol_drive}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} &= 0,\\
        \md{\rho}{t} &= Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
                e^{i(k_xx - \omega t)},\\
        \md{\vec{u}}{t} + \frac{\vec{\nabla}P}{\rho}
            + g\hat{z} &= 0.
    \end{align}
\end{subequations}

The linearized \autoref{se:vol_drive}, obtained by dropping the advective term
in the material derivative, can be solved analytically up to good accuracy. We
may first approximate the driving term using $e^{\frac{-(z - z_0)^2}{2\sigma^2}}
\approx \sqrt{2\pi \sigma^2}\delta(z - z_0)$ if $\sigma \to 0$\footnote{$\sigma$
must be large enough to be numerically resolved by the spectral code.}. This
system is solved exactly by matching the two homogeneous solutions above and
below $z_0$. We may then approximately relax the solution to $k_{0z}\sigma
\lesssim 1$: an extra factor of $e^{-\frac{(k_{0z}\sigma)^2}{2}}$ arises
compared to the $\delta$-function solution (evaluating the Fourier Transform
of $e^{-\frac{(z - z_0)^2}{2\sigma^2}}$ at $k_z = k_{0z}$), and we obtain
\begin{align}
    u_{1z}&(x, z, t) ={} \frac{Fgk_{0x}^2}{\rho_0(z_0)\omega_0^2}
        \frac{1}{2ik_z}\frac{e^{\frac{-(k_{0z}\sigma)^2}{2}}}
        {\sqrt{2\pi\sigma^2}} \times\nonumber\\
        &{}\begin{cases}
        e^{\frac{z - z_0}{2H}}e^{i(k_{0x}x + k_{0z}(z - z_0) - \omega_0 t)
            + \frac{1}{2k_zH}}
            & z > z_0\\
        e^{\frac{z - z_0}{2H}}e^{i(k_{0x}x - k_{0z}(z - z_0) - \omega_0 t)
            + \frac{1}{2k_zH}}
            & z < z_0\\
    \end{cases}.\label{eq:uz_lin}
\end{align}

@@

We nondimensionalize by taking $H = N = \rho_{ref} = 1$

where $z_B = 0.05L_z, z_T = 0.95L_z$ are the boundaries of the damping zones. A
Navier-Stokes numerical viscosity $\nu$ is used to damp high wavenumbers and
regularize the nonlinear cascade at near grid resolution: $\nu \sim 0.2
\frac{\omega}{\abs*{k_z}}\frac{L_z}{2\pi N_z}$ was found to be suitable for $N_z
= 1024$, where $k_z$ is the wavenumber of the excited linear mode.
%%%

@@

$\omega$ is chosen by inverting $\omega(k_x, k_z)$ dispersion relation for fixed
$k_x = 2\pi / L_x$ and some desired $k_z = -2\pi/H$, and $\sigma \lesssim
\frac{1}{k_z}$ is used to excite a broad band of modes including the desired
$k_z$ mode.

@@

We simulate using Dedalus (cite) a spectral numerical method. We use a Fourier
basis in the $x$ direction and a Chebyshev in the $z$ direction. Our simulation
starts with all perturbation quantities at zero, an initial condition at rest.
Our parameters are chosen to be as similar as possible to the later nonlinear
simulations (\autoref{s:nonlin}). We choose as follows:
\begin{itemize}
    \item $k_{0x}$: Astrophysical IGWs have $k_x \ll k_z$, so we choose $k_{0x}
        = \frac{2\pi}{L_x}$ the smallest permitted wavenumber permitted by
        periodic boundary conditions.

    \item We choose $\omega_0$ by choosing it to produce a desired $k_{0z}$.
        Astrophysical systems generally also exhibit $k_zH \gg 1$, where the
        stratification height is significantly larger than the vertical
        wavelength. However, in order for the simulation to both be well-damped
        at grid resolutions and be neglibily damped at leading order nonlinear
        wavelengths, we require $k_{0z} \ll \frac{1}{L_{NL}} \ll \frac{2\pi
        N_z}{L_z}$, where $L_{NL}$ denotes some characteristic length scale of
        nonlinear features. Finally, we need $L_z$ to be many $H$ to capture
        significant $e^{z/2H}$ growth and separate the linear and nonlinear
        amplitudes of $u_z$.

@@

    \item $\bar{U}_0$ agrees with \autoref{eq:u0_lin} at lower $z$ but deviates
        significantly at higher $z$. This deviation is caused by the top damping
        zone that does not conserve horizontal momentum.
\end{itemize}
We also include a qualitatively similar snapshot of the same simulation run
using the $\nu$ from the later nonlinear simulations in \autoref{fig:linear_nu}.
\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{plots/linear_nonu.png}
    \caption{Snapshot of linear simulation after reaching steady state.
        Analytical predictions of $\abs*{u_z}, \bar{U}_0, S_{px}$ are shown in
        orange. Red dotted lines indicate the onset of the damping region and
        blue dotted lines denote the driving region.}\label{fig:linear_nonu}
\end{figure}
\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{plots/linear_nu.png}
    \caption{Same as \autoref{fig:linear_nonu} but with $0.3\times $viscosity
    used for nonlinear simulations (YUBONOTE should be the same but that
    simulation mysteriously stopped without me noticing; restarting worked, so
    currently rerunning). Note the slight resulting dissipation in $S_{px}$.
    }\label{fig:linear_nu}
\end{figure}

\section{Nonlinear Regime}\label{s:nonlin}

@@

Two representative snapshots from our simulation are provided in
\autoref{fig:nl} after the critical layer has had time to form. We may note that
the critical layer, where $S_{px}$ is absorbed and $\bar{U}_0 = \bar{U}_c$,
travels downwards as predicted.
\begin{figure}[h]
    \centering
    \begin{subfigure}{0.5\textwidth}
        \centering
        \includegraphics[width=\textwidth]{plots/nl35_1.png}
        \caption{Earlier time snapshot. Legend is the same as
        \autoref{fig:linear_nonu} except $\bar{U}_c$ is marked in red on the
        third panel.}
    \end{subfigure}

@@

For the simulation in \autoref{fig:nl}, we may define $z_c = \argmax_z
\pd{\bar{U}_0}{z}$. Computing $\pd{z_c}{t}$ using the analytical flux
\autoref{eq:fpx_lin} allows comparison to \autoref{eq:zc_anal}, which we exhibit
in \autoref{fig:zc_anal}. We see overall good agreement, though some small
deviation is expected since $S_{px}$ is not perfectly conserved owing to
numerical viscosity (and YUBONOTE $S_{px}$ is misestimated?).
\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{plots/nl35_front_v.png}
    \caption{Comparison of simulated $\pd{z_c}{t}$ to analytical
    \autoref{eq:zc_anal}. Plot begins after formation of critical layer in
    simulation.}\label{fig:zc_anal}
\end{figure}

However, recalling \autoref{eq:crit_coeffs}, we only expect complete critical
layer absorption when $\mathrm{Ri} \gg \frac{1}{4}$. We may plot $\mathrm{Ri}$
over the same time interval in \autoref{fig:nl35_f_ri}. We may observe that the
Richardson number is initially decreasing but eventually levels out and
begins to increase again. This corresponds to an increasingly sharp critical
layer transition then subsequently a decreasingly sharp critical layer
transition.
\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{plots/nl35_f_ri.png}
    \caption{Plot of $\mathrm{Ri}_{\max}(t) = \max_z
    \frac{N^2}{\bar{U}_0'}(t)$ over the simulation time.}\label{fig:nl35_f_ri}
\end{figure}

We argue that the Richardson number is bounded from below by viscosity. A
repeated simulation with a larger viscosity is shown in \autoref{fig:nl1_f_ri},
where the Richardson number does not go nearly as low.
\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{plots/nl1_f_ri.png}
    \caption{Same as \autoref{fig:nl35_f_ri} but with $\sim 1.3\times$
    viscosity.}\label{fig:nl1_f_ri}
\end{figure}

Additionally, a lower-viscosity simulation shows significant features resembling
Kelvin-Helmholtz instabilities (KHI), \autoref{fig:nl_full_01}. As we have only
plotted $\bar{U}_0(z)$, the local Richardson number $N^2 / \pd{u_x}{z}$ may
exceed the plotted value and incur KHIs.
\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{plots/nl_full_01.png}
    \caption{Low viscosity simulation showing seemingly KHI-like
    features.}\label{fig:nl_full_01}
\end{figure}

This seems to suggest that our reference simulation is viscosity limited. This
is perhaps physical: a Richardson number $\sim 0.25$ incurs the KHI which saps
energy and angular momentum from the flow. It is conceivable then that our
numerical viscosity is physically motivated as a effective KHI viscosity.

@@

We denote $x \in [0, L_x], z \in [0, L_z]$ the simulation domain and $N_x, N_z$
the number of spectral modes in the respective dimensions. We perform direct
numerical simulation of \autoref{se:vol_drive} with the open-source
pseudo-spectral code Dedalus (CITE).

@@

\end{document}

>>>>

    \documentclass[twocolumn,
        nofootinbib, % unbreaks footnotes from being put in bib
        usenames, % allows access to some tikz colors
        aps,
        prd,
        dvipsnames % more colors: https://en.wikibooks.org/wiki/LaTeX/Colors
    ]{revtex4-1}% chktex 8
    \usepackage{
        amsmath,
        amssymb,
        fancyhdr, % page styling
        lastpage, % footer fanciness
        hyperref, % various links
        setspace, % line spacing
        amsthm, % newtheorem and proof environment
        mathtools, % \Aboxed for boxing inside aligns, among others
        float, % Allow [H] figure env alignment
        enumerate, % Allow custom enumerate numbering
        graphicx, % allow includegraphics with more filetypes
        wasysym, % \smiley!
        upgreek, % \upmu for \mum macro
        listings, % writing TrueType fonts and including code prettily
        tikz, % drawing things
        booktabs, % \bottomrule instead of hline apparently
        cancel % can cancel things out!
    }
    \usepackage[
        labelfont=bf, % caption names are labeled in bold
        font=scriptsize % smaller font for captions
    ]{caption}
    \usepackage[font=scriptsize]{subcaption} % subfigures

@@

In \autoref{s:equations}, we will cover the numerical setup. In \autoref{s:lin}, we
will discuss the agreement of our simulation in the low-amplitude limit with
analytical linear theory. In \autoref{s:nonlin} we will present a few reference
simulations illustrating critical layer absorption depositing horizontal angular
momentum into the fluid.

\section{Equations}\label{s:equations}

We consider an incompressible, isothermal fluid, representative of degenerate
matter in WD bulks, in 2D. We assume a barotropic equation of state as a first
approximation. As we are interested in the behavior of the wave far from the
center of the WD, we approximate the gravitational field as uniform. We model
the background density stratification as $\rho_0(x, z) = \rho_0(z) = \rho_0(z=0)
e^{-z/H}$ for some reference density $\rho_0(z=0) = \at{\rho_0(z)}_{z = 0}$.

@@

\section{Internal Gravity Waves: Linear Theory}

\subsection{Analytical Properties}

In the small perturbation limit, where flow velocities are small compared to the
characteristic space and time scales $\pd{}{t} \gg \vec{u} \cdot \vec{\nabla}$,
we linearize \autoref{se:fc_orig} by ignoring the advective components of the
material derivative, defining $\rho_1 \equiv \rho - \rho_0 \ll \rho_0 \ll
\rho_0$ and $P_1 \equiv P - P_0 \ll P_0$ deviations from equilibrium and obtain:
\begin{subequations}\label{se:lin_homo}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} &= 0,\\
        \pd{\rho_1}{t} - \frac{u_{z} \rho_0}{H} &= 0,\\
        \pd{\vec{u}}{t} + \frac{\vec{\nabla}P_1}{\rho_0}
            + \frac{\rho_1 g\hat{z}}{\rho_0}
            &= 0.
    \end{align}
\end{subequations}

It is well known that the solutions to the $F = 0$ homogeneous
\autoref{se:lin_homo} are of form\cite{drazin,sutherland0}:
\begin{subequations}\label{se:lin_sol}
    \begin{align}
        u_z\p*{x, z, t} &= Ae^{z/2H}e^{i(k_{0x}x + k_{0z}z - \omega_0 t)},\\
        \omega_0^2 &= \frac{N^2k_{0x}^2}{k_{0x}^2 + k_{0z}^2 + \frac{1}{4H^2}},
    \end{align}
\end{subequations}
up to undetermined amplitude $A$, and
\begin{equation}
    N^2 \equiv g^2\p*{\rd{\rho}{P} - \at{\pd{\rho}{P}}_{ad}}
        = \frac{g}{H},
\end{equation}
is the Brunt-V\"ais\"al\"a/buoyancy frequency for our incompressible,
exponentially stratified atmosphere.

In the weak stratification limit $k_{0z}H \gg 1$, the solution exhibits the
following characteristics:
\begin{itemize}
    \item The amplitude of the perturbations grows like $e^{z/2H}$. Thus, the
        linear approximation is violated for sufficiently large $z$.

    \item The phase and group velocities can be computed respectively:
        \begin{align}
            \vec{c}_{ph} &= \frac{Nk_{0x}}{k_{0x}^2 + k_{0z}^2 + 1/4H^2} \hat{k},\\
            \vec{c}_{g} &= \frac{N\p*{k_{0x}k_{0z}\hat{x}
                - \p*{k_{0z}^2 + \frac{1}{4H^2}}\hat{z}}}
                {\p*{k_{0x}^2 + k_{0z}^2 + \frac{1}{4H^2}}^{3/2}} \nonumber\\
                &\approx \frac{Nk_{0z}}{k_{0x}^2 + k_{0z}^2}\p*{k_{0x}\hat{x} -
                k_{0z}\hat{z}}.
        \end{align}
        We recover the usual result $\vec{c}_{ph} \cdot \vec{c}_g = 0$
        (e.g.\ \cite{drazin,sutherland1}). Note that an IGW transporting energy
        and momentum upwards $c_{g, z} > 0$ has $k_{0z} < 0$.

    \item The time-averaged total $x$-momentum flux in the $\hat{z}$ direction
        can be computed. Since the linear solution is separable as $f(x, z, t) =
        f(z)e^{i(k_{0x}x - \omega_0 t)}$, $x$ averaging and time averaging are
        equivalent, so we may write ($\ev*{\dots}_x$ denotes $x$-averaging)
        \begin{align}
            S_{px, z} &\equiv \ev*{\rho u_x u_z}_x \equiv
                \frac{1}{L_x}\int_0^{L_x}\limits \rho u_x u_z\;\mathrm{d}x
                    \nonumber\\
                &\approx -\frac{A^2}{2}\rho_0(z=0)\frac{k_{0z}}{k_{0x}}.
                    \label{eq:fpx_lin}
        \end{align}

    \item Finally, it may be noted that horizontally averaged horizontal
        momentum $p_x = \ev*{\rho u_x}_x$ obeys $\pd{p_x}{t} + \pd{S_{px}}{z} =
        0$. Since flux is transported on the group velocity
        timescale\cite{sutherland0}, this simplifies to the known expression for
        the wave-induced horizontal mean flow\cite{sutherland0,sutherland1}
        \begin{equation}
            \bar{U}_0 \equiv \ev*{u_x}_x = \frac{\ev*{u_xu_z}_x}{c_{gz}}
        \end{equation}
        We've denoted $c_{gz} = \at{\pd{\omega}{k_{z}}}_{k_{z0}}$ the vertical group
        velocity. This may also be rewritten
        \begin{equation}
            \bar{U}_0 = -\frac{A^2}{2}e^{\frac{z - z_0}{H}}
                \frac{k_{0z}}{k_{0x}}c_{gz}, \label{eq:u0_lin}
        \end{equation}
        where $c_{gz} = -\frac{Nk_{0x}k_{0z}}{\p*{k_{0x}^2 + k_{0z}^2 +
        \frac{1}{4H^2}}^{3/2}}$.
\end{itemize}

@@

As not to interfere with the incompressibility constraint, we force the system
on the density equation, replacing \autoref{eq:fc_density} with
\begin{equation}
    \md{\rho}{t} = Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
            \cos \p*{k_{0x}x - \omega_0 t}.\label{eq:vol_drive}
\end{equation}
Using a narrow Gaussian profile excites a broad $k_z$ wavenumber spectrum, and
only the $k_{0z}$ satisfying the IGW's dispersion relation for the given
$k_{0x}, \omega_0(k_{0x}, k_{0z})$ will propagate.

The linearized \autoref{eq:vol_drive}, obtained by dropping the advective term
in the material derivative, can be solved analytically up to good accuracy. We
may first approximate the driving term using $e^{\frac{-(z - z_0)^2}{2\sigma^2}}
\approx \sqrt{2\pi \sigma^2}\delta(z - z_0)$ if $\sigma \to 0$\footnote{$\sigma$
must be large enough to be numerically resolved by the spectral code.}. This
system is solved exactly by matching the two homogeneous solutions above and
below $z_0$. We may then approximately relax the solution to $k_{0z}\sigma
\lesssim 1$: an extra factor of $e^{-\frac{(k_{0z}\sigma)^2}{2}}$ arises
compared to the $\delta$-function solution (evaluating the Fourier Transform
of $e^{-\frac{(z - z_0)^2}{2\sigma^2}}$ at $k_z = k_{0z}$), and we obtain
\begin{align}
    u_{1z}&(x, z, t) ={} \frac{Fgk_{0x}^2}{\rho_0(z_0)\omega_0^2}
        \frac{1}{2ik_{0z}}\frac{e^{\frac{-(k_{0z}\sigma)^2}{2}}}
        {\sqrt{2\pi\sigma^2}} \times\nonumber\\
        &{}\begin{cases}
        e^{\frac{z - z_0}{2H}}e^{i(k_{0x}x + k_{0z}(z - z_0) - \omega_0 t)
            + \frac{1}{2k_{0z}H}}
            & z > z_0\\
        e^{\frac{z - z_0}{2H}}e^{i(k_{0x}x - k_{0z}(z - z_0) - \omega_0 t)
            + \frac{1}{2k_{0z}H}}
            & z < z_0\\
    \end{cases}.\label{eq:uz_lin}
\end{align}
The $z > z_0$ solution models an upwards-propagating IGW wavetrain inbound on
the simulation domain from below. The solution for $u_{1x}(x, z, t)$ moreover is
just $u_{1x} \approx \frac{k_{0z} u_{1z}}{k_{0x}}e^{\frac{1}{2k_{0z}H}}$.

\section{Numerical Simulation}

@@

We nondimensionalize by taking $H = N = \rho_0(z=0) = 1$ in
\autoref{se:fc_orig}.

where $z_B = 0.05L_z, z_T = 0.95L_z$ are the boundaries of the damping zones. A
Navier-Stokes numerical viscosity $\nu$ is used to damp high wavenumbers and
regularize the nonlinear cascade at near grid resolution: $\nu \sim 0.2
\frac{\omega}{\abs*{k_{0z}}}\frac{L_z}{2\pi N_z}$ was found to be suitable for
$N_z = 1024$, where $k_{0z}$ is the wavenumber of the excited linear mode.
%%%

@@

$\omega$ is chosen by inverting $\omega(k_{0x}, k_{0z})$ dispersion relation for fixed
$k_{0x} = 2\pi / L_x$ and some desired $k_{0z} = -2\pi/H$, and $\sigma \lesssim
\frac{1}{k_{0z}}$ is used to excite a broad band of $k_z$ modes including the
desired $k_{0z}$ mode.

@@

% Nonlinear: Critical layer absorption (viscous = Hazel, nonlin = Winters),
% wave-induced mean flow, horizontal momentum flux

We simulate using Dedalus\cite{dedalus} a spectral numerical method. We use a
Fourier basis in the $x$ direction and a Chebyshev in the $z$ direction. Our
simulation starts with all perturbation quantities at zero, an initial condition
at rest. Our parameters are chosen to be as similar as possible to the later
nonlinear simulations (\autoref{s:nonlin}). We choose as follows:
\begin{itemize}
    \item $k_{0x}$: Astrophysical IGWs have $k_{0x} \ll k_{0z}$, so we choose $k_{0x}
        = \frac{2\pi}{L_x}$ the smallest permitted wavenumber permitted by
        periodic boundary conditions.

    \item We choose $\omega_0$ by choosing it to produce a desired $k_{0z}$.
        Astrophysical systems generally also exhibit $k_{0z}H \gg 1$, where the
        stratification height is significantly larger than the vertical
        wavelength. However, in order for the simulation to both be well-damped
        at grid resolutions and be neglibily damped at leading order nonlinear
        wavelengths, we require $k_{0z} \ll \frac{1}{L_{NL}} \ll \frac{2\pi
        N_z}{L_z}$, where $L_{NL}$ denotes some characteristic length scale of
        nonlinear features. Finally, we need $L_z$ to be many $H$ to capture
        significant $e^{z/2H}$ growth and separate the linear and nonlinear
        amplitudes of $u_z$.

@@

    \item $\bar{U}_0$ agrees with \autoref{eq:u0_lin} at lower $z$ but deviates
        significantly at higher $z$. This deviation is caused by the top damping
        zone that does not conserve horizontal momentum.
\end{itemize}
We also include a qualitatively similar snapshot of the same simulation run
using the $\nu$ from the later nonlinear simulations in \autoref{fig:linear_nu}.
\begin{figure}[t]
    \centering
    \includegraphics[width=0.5\textwidth]{plots/linear_nonu.png}
    \caption{Snapshot of linear simulation after reaching steady state.
        Analytical predictions of $\abs*{u_z}, \bar{U}_0, S_{px}$ are shown in
        orange. Red dotted lines indicate the onset of the damping region and
        blue dotted lines denote the driving region.}\label{fig:linear_nonu}
\end{figure}
\begin{figure}[t]
    \centering
    \includegraphics[width=0.5\textwidth]{plots/linear_nu.png}
    \caption{Same as \autoref{fig:linear_nonu} but with $0.3\times $viscosity
    used for nonlinear simulations (YUBONOTE should be the same but that
    simulation mysteriously stopped without me noticing; restarting worked, so
    currently rerunning). Note the slight resulting dissipation in $S_{px}$.
    }\label{fig:linear_nu}
\end{figure}

\subsection{Nonlinear Regime}\label{s:nonlin}

@@

Two representative snapshots from our simulation are provided in
\autoref{fig:nl} after the critical layer has had time to form. We may note that
the critical layer, where $S_{px}$ is absorbed and $\bar{U}_0 = \bar{U}_c$,
travels downwards as predicted.
\begin{figure}[t]
    \centering
    \begin{subfigure}{0.5\textwidth}
        \centering
        \includegraphics[width=\textwidth]{plots/nl35_1.png}
        \caption{Earlier time snapshot. Legend is the same as
        \autoref{fig:linear_nonu} except $\bar{U}_c$ is marked in red on the
        third panel.}
    \end{subfigure}

@@

For the simulation in \autoref{fig:nl}, we may define $z_c = \argmax_z
\pd{\bar{U}_0}{z}$. Computing $\pd{z_c}{t}$ using the analytical flux
\autoref{eq:fpx_lin} allows comparison to \autoref{eq:zc_anal}, which we exhibit
in \autoref{fig:zc_anal}. We see overall good agreement, though some small
deviation is expected since $S_{px}$ is not perfectly conserved owing to
numerical viscosity (and YUBONOTE $S_{px}$ is misestimated?).
\begin{figure}[t]
    \centering
    \includegraphics[width=0.5\textwidth]{plots/nl35_front_v.png}
    \caption{Comparison of simulated $\pd{z_c}{t}$ to analytical
    \autoref{eq:zc_anal}. Plot begins after formation of critical layer in
    simulation.}\label{fig:zc_anal}
\end{figure}

However, recalling \autoref{eq:crit_coeffs}, we only expect complete critical
layer absorption when $\mathrm{Ri} \gg \frac{1}{4}$. We may plot $\mathrm{Ri}$
over the same time interval in \autoref{fig:nl35_f_ri}. We may observe that the
Richardson number is initially decreasing but eventually levels out and
begins to increase again. This corresponds to an increasingly sharp critical
layer transition then subsequently a decreasingly sharp critical layer
transition.
\begin{figure}[t]
    \centering
    \includegraphics[width=0.5\textwidth]{plots/nl35_f_ri.png}
    \caption{Plot of $\mathrm{Ri}_{\max}(t) = \max_z
    \frac{N^2}{\bar{U}_0'}(t)$ over the simulation time.}\label{fig:nl35_f_ri}
\end{figure}

We argue that the Richardson number is bounded from below by viscosity. A
repeated simulation with a larger viscosity is shown in \autoref{fig:nl1_f_ri},
where the Richardson number does not go nearly as low.
\begin{figure}[t]
    \centering
    \includegraphics[width=0.5\textwidth]{plots/nl1_f_ri.png}
    \caption{Same as \autoref{fig:nl35_f_ri} but with $\sim 1.3\times$
    viscosity.}\label{fig:nl1_f_ri}
\end{figure}

@@

We denote $x \in [0, L_x], z \in [0, L_z]$ the simulation domain and $N_x, N_z$
the number of spectral modes in the respective dimensions. We perform direct
numerical simulation of \autoref{eq:vol_drive} with the open-source
pseudo-spectral code Dedalus (CITE).

@@

\bibliographystyle{apsrev4-1} % chktex 8
\bibliography{paper}
\end{document}
